The spatial patterning of Middle Paleolithic human settlement in westernmost Iberia

### 2020-11-04 README.Rmd
submission

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "The spatial patterning of Middle Paleolithic human settlement in Western Iberia"

@@

# The spatial patterning of Middle Paleolithic human settlement in Western Iberia

>>>>

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "The spatial patterning of Middle Paleolithic human settlement in westernmost Iberia"

@@

# The spatial patterning of Middle Paleolithic human settlement in westernmost Iberia

### 2020-11-04 analysis/paper/paper.Rmd
submission

---
title: "The spatial patterning of Middle Paleolithic human settlement in Western Iberia"
author:
  - João Cascalheira
  - Célia Gonçalves
  - Daniela Maio
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/elsevier-harvard.csl" # Insert path for the bib-style
abstract: |
  Currently available data on the Pleistocene human occupation of the westernmost territories of Iberia attests the presence of Middle Paleolithic industries from c. 240 ka cal BP until c. 37 ka cal BP. Previous studies focusing on this timeframe have suggested that Middle Paleolithic populations were highly mobile and predominately utilized locally available raw materials, with many of cave and open-air sites being located near fluvial settings. Other than these observations, no specific studies have focused on exploring the factors influencing human site location choice during that time range. Employing statistical and GIS approaches, this paper provides an initial assessment of spatial patterning in human settlement during the Middle Paleolithic of westernmost Iberia. Results show that sites’ location are biased towards lower elevations and riverine settings and suggest that distance to rivers might have had an impact on the diversity and specific types of lithic raw materials used at each site. These results help to shed light on the particularities of Neanderthal adaptations in a region regarded as refugium during periods of unfavorable climate during the Middle Paleolithic.
keywords: |
  GIS; Settlement patterns; Middle Paleolithic; Western Iberia; Portugal
---

>>>>

---
title: "The spatial patterning of Middle Paleolithic human settlement in westernmost Iberia"
author:
  - João Cascalheira
  - Célia Gonçalves
  - Daniela Maio
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/elsevier-harvard.csl" # Insert path for the bib-style
abstract: |
  Currently available data on the Pleistocene human occupation of the westernmost territories of Iberia attests the presence of Middle Paleolithic industries from c. 240 ka cal BP until c. 37 ka cal BP. Previous studies focusing on this timeframe have suggested that Middle Paleolithic populations were highly mobile and predominately utilized locally available raw materials, with many of cave and open-air sites being located near fluvial settings. Other than these observations, no specific studies have focused on exploring the factors influencing human site location choice during that time range. Employing statistical and GIS approaches, this paper provides an initial assessment of spatial patterning in human settlement during the Middle Paleolithic of westernmost Iberia. Results show that sites’ location are biased towards lower elevations and riverine settings and suggest that distance to rivers might have had an impact on the diversity and specific types of lithic raw materials used at each site. These results help to shed light on the particularities of Neanderthal adaptations in a region regarded as refugium during periods of unfavorable climate during the Middle Paleolithic.
keywords: |
  GIS; Settlement patterns; Middle Paleolithic; Western Iberia; Portugal
---
